### 2022-11-28 dissertation/chapters/07-methodology.tex
✨ (final-review): Chapter 7 Done

Each empirical test starts with a question to be answered. A hypothesis is formulated for the outcome of the empirical test. The empirical tests are then designed around the implementation of the components under evaluation. Each empirical test defines the configuration of elements and generally include a set of parameters that are altered between experiments. Each experiment is evaluated by means of a performance measurement. In the context of training \acp{FFNN}, the underlying model is trained across a number of datasets. Each dataset is split into a training set comprising of 80\% of the data, and a test set comprising of 20\% of the data. The training set is used to train the model, while the test set is used to evaluate the model's performance on unseen data. Training epochs are split into mini-batch iterations with a specified mini-batch size per dataset. Evaluation takes place at every mini-batch step. Due to the stochastic nature of the experiments, each experiment is repeated over a number of runs to provide sufficient sample for statistical certainty about the outcome of the results. Each run contains a different random seed, such that each run is unique. The evaluation data forms the results of the empirical test. These results are analysed for statistical significance, from which findings and conclusions are then made. The null hypothesis is then either rejected or verified based on these findings.

Details around each of these elements are provided in the following sections and lead with discussions around the datasets that are used.

@@

All models that are trained in this dissertation follow implementations of shallow \acp{FFNN}, meaning that they only have one hidden layer. The architecture of a model is dependent on the dataset being trained on, the type of optimisation problem, the number of input dimensions and the number of output dimensions. The number of hidden layers were determined empirically. Weights are initialised by means of \index{Glorot uniform sampling}Glorot uniform sampling~\cite{ref:glorot:2010}. The models and their configuration, as it is used for each dataset, are given in Table \ref{tab:methodology:models:configurations}.

@@

This section provides the details of the low-level \index{heuristic}heuristics that are used in the empirical process. Each of these low-level \index{heuristic}heuristics are implemented as standalone \index{heuristic}heuristics and are also included in the \index{heuristic pool}heuristic pool of the \acs{BHH}. Table \ref{tab:methodology:heuristics} contains a list of all the standalone, low-level \index{heuristic}heuristics that are used as well as their hyper-parameter configuration.

% Table generated by Excel2LaTeX from sheet 'BHH Baseline'
\begin{table}[htbp]
      \centering
      \caption{Low-level \index{heuristic}heuristics and their hyper-parameter configuration.}
      \label{tab:methodology:heuristics}%
      \par\bigskip
      \resizebox{0.7\textwidth}{!}{

@@

This section provides the performance measures that are used to evaluate the different experimental runs during the empirical process. Chapter \ref{chap:anns} provided a number of loss functions. These loss functions are used to measure the performance of the \acp{FFNN} being trained. In this dissertation, \acf{BinXE} is used for classification problems with two classes and \acf{SparseCatXE} is used for classification problems with more than two classes. For the classifications problems accuracy is also measured. Accuracy refers to the proportion of correct classifications. For regression problems, the \acf{RMSE} is used as a performance metric.

@@

This section provides the stopping conditions that are used during the empirical process. For this dissertation, the \textit{maximum epoch} stopping condition is used, where a fixed, maximum number of training steps and epochs are executed and training is not haulted until the maximum number of epochs have been reached. Early stopping is a technique where training is stopped when the \index{heuristic}heuristic is not able to improve the performance of the \acs{FFNN} for a number of steps. If the performance has not improved, the training is halted and the last known best model weights are restored. Early stopping is not implemented for the empirical process in this dissertation. This design decision is made so that the behaviour of the \acs{BHH} can be studied beyond the point of optimal results.

@@

The behavioural case study analyses the behaviour of the \acs{BHH} baseline configuration as it relates to an example problem dataset, across 30 independent runs. The example problem dataset is the iris dataset, presented in Table \ref{tab:methodology:datasets:classification}. The behavioural case study is meant to provide an introductory analysis of the inner workings of the \acs{BHH} and includes analysis of the selection mechanism, as well as the perturbative nature of the \acs{BHH}. The behavioural case study is not meant to be statistically analysed, but to rather provide a detailed visual analysis of a number of hand-picked example runs, to determine if the \acs{BHH} is behaving as expected. From these observations, it is determined if the \acs{BHH} is learning and that selection is indeed biasing towards better performance. These observations also provide an opportunity to observe the outcome of the perturbative nature of the \acs{BHH}, which includes proxied \index{heuristic}heuristic update step operations.

@@

Various experimental groups and configurations have been presented in Sections \ref{sec:methodology:experiments:case_study}, \ref{sec:methodology:experiments:standalone_optimisers} and \ref{sec:methodology:experiments:bhh_variants}. The results from each of these experimental groups are statistically analysed. To ensure that there are sufficient samples for statistical analysis, each experiment and configuration is trained for 30 epochs and is repeated over 30 runs, for each of the 15 datasets. The experimental results contain performance evaluation data for the training and testing datasets, and includes loss and accuracy measurements. Statistical analysis is executed on the results from the testing datasets. An average rank is calculated across all 30 runs, for all experimental groups and configurations, at each step, for every epoch of training. Each run is executed using a unique random seed, such that each run is identical in its setup and configuration, but unique between runs.

>>>>

Each empirical test starts with a question to be answered. A hypothesis is formulated for the outcome of the empirical test. The empirical tests are then designed around the implementation of the components under evaluation. Each empirical test defines the configuration of elements and generally include a set of parameters that are altered between experiments. Each experiment is evaluated by means of a performance measurement. In the context of training \acp{FFNN}, the underlying model is trained across a number of datasets. Each dataset is split into a training set comprising of 80\% of the data, and a test set comprising of 20\% of the data. The training set is used to train the model, while the test set is used to evaluate the model's performance on unseen data. Training epochs are split into mini-batch iterations with a specified mini-batch size per dataset. Evaluation takes place at every mini-batch step. Due to the stochastic nature of the experiments, each experiment is repeated over 30 independent runs to provide sufficient sample for statistical certainty about the outcome of the results. Each run contains a different random seed, such that each run is unique. The evaluation data forms the results of the empirical test. These results are analysed for statistical significance, from which findings and conclusions are then made. The null hypothesis is then either rejected or accepted based on these findings.

Details around each of these elements are provided in the following sections.

@@

All models that are trained in this dissertation follow implementations of shallow \acp{FFNN}, meaning that they only have one hidden layer. The architecture of a model is dependent on the dataset being trained on, the type of optimisation problem, the number of input dimensions and the number of output dimensions. The number of hidden layers used were determined empirically. Weights are initialised by means of \index{Glorot uniform sampling}Glorot uniform sampling~\cite{ref:glorot:2010}. The models and their configuration, as it is used for each dataset, are given in Table \ref{tab:methodology:models:configurations}.

@@

This section provides the details of the low-level \index{heuristic}heuristics that are used in the empirical process. Each of these low-level \index{heuristic}heuristics are implemented as standalone \index{heuristic}heuristics and are also included in the \index{heuristic pool}heuristic pool of the \acs{BHH}. Table \ref{tab:methodology:heuristics} contains a list of all the standalone, low-level \index{heuristic}heuristics that are used as well as their hyper-parameter configurations.

% Table generated by Excel2LaTeX from sheet 'BHH Baseline'
\begin{table}[htbp]
      \centering
      \caption{Low-level \index{heuristic}heuristics and their hyper-parameter configurations.}
      \label{tab:methodology:heuristics}%
      \par\bigskip
      \resizebox{0.7\textwidth}{!}{

@@

This section provides the performance measures that are used to evaluate the different experimental runs during the empirical process. Chapter \ref{chap:anns} provided a number of loss functions. These loss functions are used to measure the performance of the \acp{FFNN} being trained. In this dissertation, \acs{BinXE} is used for classification problems with two classes and \acs{SparseCatXE} is used for classification problems with more than two classes. For the classification problems, accuracy is also measured. Accuracy refers to the proportion of correct classifications. For regression problems, the \acs{RMSE} is used as a performance metric.

@@

This section provides the stopping conditions that are used during the empirical process. For this dissertation, the \textit{maximum epoch} stopping condition is used, where a fixed, maximum number of training steps and epochs are executed and training is not haulted until the maximum number of epochs have been reached. \acp{FFNN} are trained for a maximum of 30 epochs. Early stopping is a technique where training is stopped when the \index{heuristic}heuristic is not able to improve the performance of the \acs{FFNN} for a number of steps. If the performance has not improved, the training is halted and the last known best model weights are restored. Early stopping is not implemented for the empirical process in this dissertation. This design decision is made so that the behaviour of the \acs{BHH} can be studied beyond the point of optimal results.

@@

The behavioural case study analyses the behaviour of the \acs{BHH} baseline configuration as it relates to an example problem dataset, across 30 independent runs. The example problem dataset is the iris dataset, presented in Table \ref{tab:methodology:datasets:classification}. The behavioural case study is meant to provide an introductory analysis of the inner workings of the \acs{BHH} and includes analysis of the selection mechanism, as well as the perturbative nature of the \acs{BHH}. The behavioural case study is used to determine if the \acs{BHH} is learning and that selection is indeed biasing towards better performance. These observations also provide an opportunity to observe the outcome of the perturbative nature of the \acs{BHH}, which includes proxied \index{heuristic}heuristic update step operations.

@@

Various experimental groups and configurations have been presented in Sections \ref{sec:methodology:experiments:case_study}, \ref{sec:methodology:experiments:standalone_optimisers} and \ref{sec:methodology:experiments:bhh_variants}. The results from each of these experimental groups are statistically analysed. To ensure that there are sufficient samples for statistical analysis, each experiment and configuration is trained for 30 epochs and is repeated over 30 runs, for each of the fourteen datasets. The experimental results contain performance evaluation data for the training and testing datasets, and includes loss and accuracy measurements. Statistical analysis is executed on the results from the testing datasets. An average rank is calculated across all 30 runs, for all experimental groups and configurations, at each step, for every epoch of training. Each run is executed using a unique random seed, such that each run is identical in its setup and configuration, but unique between runs.
